curatedMetagenomicData

### 2019-12-03 combodata_NS.Rmd
updated

@@

@@

# this defines the colours for the graphs. Uses the wes palettes. Look at https://github.com/karthik/wesanderson for the different themes
glob_theme <- "FantasticFox1"

my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category | mnd, data=data)
}

@@

  # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  print("The basic aloha diversity analysis has been completed")
  # This country versus nationality
  p1<-   ggviolin(pseq.meta, x = "country", y = "Shannon",
                  add = "boxplot", fill = "country", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  print("First graph has completed")
  # the number in the palettes must equal the number of groups in bmi.pairs

@@

  print("Third graph has completed")
  ggsave(paste("/Users/Nik/Alpha_diversity_",title_text,".png", sep = "" ), device = "png")
  #ggsave("/Users/Nik/Alpha_diversity_curate.png")
}

@@

  # This add the facets but I do not think we need that any more. 
  #p3 <- p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
  ggsave(paste("Beta_diversity_",title_text,"_NMDS.png", sep = "" ), device = "png")
    # MDS
   pseq.mds <<- ordinate(pseq.subset, "MDS", "bray")
   p4 <- plot_ordination(pseq.subset, pseq.mds, type = "sample", color = "mnd")
 #plot(p1 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence MDS"))

    p5 <- p4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND Prevalence") + theme(legend.position = "none")
    ggsave(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), device = "png")

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND Prevalence") + theme(legend.position = "bottom")
    ggsave(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), device = "png")
       #this just generates the titles for the output
    title <- ggdraw() + 
    draw_label(
    title_text,
    fontface = 'bold',
    x = 0,
    hjust = 0
    ) +
    theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
    )

    p10 <<-plot_grid(title,p3, p5,p8, ncol = 1, labels = c('','A', 'B','C'),rel_heights = c(0.3, 1.5,1.5,2))
    ggsave(paste("Beta_diversity_ALL_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")
}

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  ggsave(paste("PERMANOVA_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

pre_perma_d(pseq.subset) 
perma_d(p.med,p.high,"RLWT_Curatefinal_highVmed")
p

@@

pre_perma_d(atlas.pseq) 
perma_d(p.med,p.high,"RLWT_Atlasfinal_highVmed")
p

>>>>

setwd("~/globalmicrobiome")

@@

#drops unused levels
c.mnd.atlas <- droplevels(c.mnd.atlas,c("ISL","HUN","EST", "SVK", "LUX", "CAN", "AUT", "BGD","FJI") )

sample_data(pseq)$nationality = c.mnd.atlas

@@

# this defines the colours for the graphs. Uses the wes palettes. Look at https://github.com/karthik/wesanderson for the different themes
glob_theme <- "FantasticFox1"
setwd("~/globalmicrobiome") 
my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category | mnd, data=data)
}

@@

  # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  print("The basic alpha diversity analysis has been completed")
  # This country versus nationality
  p1<-   ggviolin(pseq.meta, x = "nationality", y = "Shannon",
                  add = "boxplot", fill = "nationality", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  print("First graph has completed")
  # the number in the palettes must equal the number of groups in bmi.pairs

@@

  print("Third graph has completed")
  save_plot(paste(title_text,".png", sep = ""),p3,base_height = 11.69 )
  #ggsave("/Users/Nik/Alpha_diversity_curate.png")
}

@@

  # This add the facets but I do not think we need that any more. 
  #p3 <- p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
  save_plot(paste("Beta_diversity_",title_text,"_NMDS.png", sep = "" ), p3)
    # MDS
   pseq.mds <<- ordinate(pseq.subset, "MDS", "bray")
   p4 <- plot_ordination(pseq.subset, pseq.mds, type = "sample", color = "mnd")
 #plot(p1 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence MDS"))

    p5 <- p4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND Prevalence") + theme(legend.position = "none")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p5)

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND Prevalence") + theme(legend.position = "bottom")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p8)
       #this just generates the titles for the output
    title <- ggdraw() + 
    draw_label(
    title_text,
    fontface = 'bold',
    x = 0,
    hjust = 0
    ) +
    theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
    )

    p10 <<-plot_grid(title,p3, p5,p8, ncol = 1, labels = c('','A', 'B','C'),rel_heights = c(0.3, 1.5,1.5,2))
    save_plot(paste("Beta_diversity_ALL_",title_text,".png", sep = "" ), p10)
}

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

pre_perma_d(pseq.subset) 
perma_d(p.med,p.high,"RLWT_Curatefinal_highVmed")

@@

pre_perma_d(atlas.pseq) 
perma_d(p.med,p.high,"RLWT_Atlasfinal_highVmed")

#   not working from here !!!!!!
